Software for the analysis and visualization of deep mutational scanning data

### 2015-03-24 paper/BMC_Bioinformatics_2015/ResponseToReviewers.tex

\begin{response}
This is a \emph{hugely} important point! The first paragraph of the subsection in the \textsl{Background} entitled \textsl{The goal: inferring site-specific amino-acid preferences} now discusses this issue in detail. Essentially, assuming the length $L$ of the gene is $\gg 1$ (which is the case for any reasonable gene), then if the mutant library contains a Poisson distribution of the number of mutations per gene (as is often the case), the problem noted by the reviewer will not arise in practice. The reason is that for such genes, knowing that there is a mutation at site $r$ does not appreciably change the marginal distribution of the number of mutations at other sites (it only changes it by a factor of $\frac{L - 1}{L}$, which in general will be very small). On the other hand, if the mutant library is created such that it contains genes with exactly one mutation per gene (which is now sometimes done), then the problem noted by the reviewer will arise even for realistic genes with $L \gg 1$. The subsection \textsl{The goal: inferring site-specific amino-acid preferences} discusses this issue, and how the problem can be corrected by appropriately defining the frequencies for the wildtype character.
\end{response}

3. Given the substantial increase in runtime and computational complexity of the Bayesian/MCMC approach relative to the simple ratio approach (hours vs fraction of a second), it might be helpful to provide a bit more intuition about what is gained. Is it simply a ``regularization'' where high ratios obtained from low counts are pushed back towards the prior? If so, does the framework really provide more than what could be obtained with a much simpler regularization heuristic?

\begin{response}
The reviewer asks the eminently reasonable question, ``Is the computational complexity of the Bayesian inference worth it?'' The revised manuscript includes a new subsection entitled \textsl{Is the Bayesian inference worthwhile?} directly after the description of the preference inference on simulated data. The honest answer to this question is equivocal -- the Bayesian inference appears to always perform as well or better than simple ratio calculations, but the improvements are non-negligible only in some situations. Hopefully the discussion in this new subsection provides a fair discussion of this topic.

Note that even when ratio estimation (which is also implemented in \texttt{dms\_tools} via the \texttt{--ratio\_estimation} option) is just as good, the visualization capabilities of \texttt{dms\_tools} will still be useful.
\end{response}

@@

4. It’s not clear whether \texttt{dms\_inferprefs} outputs the probability of $\phi_{r,x} < 1$ (or $\phi_{r,x} \ne 0$). This is likely to be a useful feature.

5. It's not clear whether the probability values for $\Delta\pi_{r,x} > 0$ or $< 0$ for \texttt{dms\_inferdiffprefs} should be interpreted as nominal or corrected. Providing false discovery rate estimates might be helpful for interpretation (also applies to 4.).

\begin{response}
The posterior probabilities $\Pr\left(\Delta\pi_{r,x} > 0 \mid \mbox{data}\right)$ reported by \texttt{dms\_inferdiffprefs} are \emph{not} corrected. However, note that these are Bayesian posterior probabilities, not classical $P$-values. By default, \texttt{dms\_inferdiffprefs} places a prior over the $\Delta\pi_{r,x}$ values that biases them towards zero (see Equation 31 of the manuscript). Therefore, it is not clear that these posterior probabilities need to be corrected -- there are arguments for why Bayesian posterior probabilities inferred with appropriate priors do not need to be corrected (see \url{http://www.stat.columbia.edu/~gelman/research/published/multiple2f.pdf}). Rather than discuss these complex and incompletely resolved statistical questions, we have simply added the following sentence:
\begin{quote}
In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (these posterior probabilities are \emph{not} corrected to account for the fact that multiple sites are typically being examined, although be default the inferences are made using the regularizing prior in Equation 31).
\end{quote}
\end{response}

6. The function of \texttt{dms\_merge} isn't explained in the paper. Is it simply averaging the two replicate profiles? Could additional statistical power be gained by explicitly modeling replicates within the Bayesian framework?

### 2015-03-24 paper/BMC_Bioinformatics_2015/bmc_bioinformatics_article.tex
Responded to reviewer 1 comment 3

>>>>

\subsubsection*{Is the Bayesian inference worthwhile?}
The foregoing sections explain why the Bayesian inference of preferences implemented in \texttt{dms\_tools} is conceptually preferable to estimating mutational effects via direct calculation of enrichment ratios using Equation \ref{eq:enrichmentcounts}. However, do the practical benefits of this Bayesian inference justify its increased complexity? The simulations in the previous section show that the Bayesian inference is more accurate, but in the absence of background errors (Figure \ref{fig:PrefCorrelations}A) the magnitude of the improvement becomes negligible once the mutations counts per site $N \overline{\mu}$ start to exceed $\sim10^3$. When there is a need to correct for background errors (Figure \ref{fig:PrefCorrelations}B), meaningful benefits of the Bayesian inference over enrichment ratios extend to somewhat higher sequencing depths. Overall, it appears that Bayesian inference will always perform as well or better than enrichment ratios. But the tangible benefit becomes negligible at high sequencing depth. In that case, the user will have to decide if the increased computational runtime and complexity of the Bayesian inference is worth a small improvement in accuracy. Simpler enrichment-ratio calculations can be performed using the \texttt{--ratio\_estimation} option of \texttt{dms\_inferprefs} or using an alternative program such as \texttt{Enrich}~\cite{fowler2011enrich}. When applying enrichment-ratio calculations to data where some mutations have low counts, it is important to include pseudocounts (denoted by $\mathcal{P}$ in Equation \ref{eq:enrichmentcounts}) as a form of regularization to avoid estimating excessively high or low ratios at sites with limited counts.
